---
title: "DD_fmri"
author: "Sophie Fromm"
date: "14 6 2021"
output: word_document
theme: yeti
#bibliography: lib_DD.bib
---

```{r setup_loadData, include=FALSE}
rm(list=ls())

## Preparations - Load  packages and data 
library(readxl)
library(tidyverse)
library(ggpubr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(kableExtra)
library(emmeans)
library(lsmeans)
library(afex)
library(Rfit)
library(janitor)
library(Hmisc)
library(DescTools)
library(rstatix)
library(car)
library(PerformanceAnalytics)

#round to 2 and pvalues to 3 digits after comma
options(scipen = 1, digits = 2)
f_pvalue = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 3)
  if (p.value <.001) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 3)))
  }
}
```

```{r LoadData, include=FALSE}
#Load data n = 86
DD_data_all <- readxl::read_excel("RDissociateDelusion_data_cleanR.xlsx")

#Create groups
DD_data_all$cat_group <- as.integer(as.factor(DD_data_all$group_DD))
mapping <- c("1" = "HC", "2" = "HC+D", "3" = "SZ","4" = "SZ+D")
DD_data_all$cat_group<- mapping[DD_data_all$cat_group]

#factorize Illness and Delusion
DD_data_all$fIllness <-factor(DD_data_all$Illness)
DD_data_all$fDelusion <-factor(DD_data_all$Delusion)
```

```{r selectSamples, include=FALSE}
#aborted trajectories results in n = 82
DD_data<-filter(DD_data_all,Code!="sz_10_014" & Code!="sz_20_123" & Code!="sz_20_136" & Code!= "sz_20_196")

#Nonfitter results in n = 66
DD_data_fitter<-filter(DD_data, Code!="sz_20_178" & Code!='sz_20_194' & Code!='sz_20_195' & Code!='sz_00_022'& Code!="sz_00_083" & Code!="sz_10_020" &  Code!='sz_10_033'& Code!='sz_10_036'& Code!="sz_10_050"  & Code!='sz_20_023'  & Code!="sz_20_082" & Code!="sz_20_087" & Code!='sz_20_092'& Code!='sz_20_120' & Code!="sz_20_154" &  Code!='sz_20_158')
#& Code!='sz_20_035' & Code!="sz_20_124" & Code!='sz_20_108' & Code!='sz_00_102')Code!='sz_20_182'  &

#long format
DD_data_long<-gather(DD_data, parameter_type, parameter_value,  M6_kappa:M6_beta)
DD_data_long_fitter <-gather(DD_data_fitter, parameter_type, parameter_value,  M6_kappa:M6_beta)
```

### Participants
Our sample (n=`r nrow(DD_data)`, `r mean(DD_data$Alter)` ± `r sd(DD_data$Alter)`y, `r sum(DD_data$Geschlecht == "w")` females) comprised healthy participants (HC) with high (n= `r sum(DD_data$cat_group == "HC+D")` and without delusional ideation (n= `r sum(DD_data$cat_group == "HC")`), participants with diagnosed schizophrenia (SZ) with (n= `r sum(DD_data$cat_group == "SZ+D")`) and without delusions (n= `r sum(DD_data$cat_group == "SZ")`). Delusional ideation was determined based on the PDI-total score [@Peters2001] (total PDI < 7 for D- and total PDI > 9 for D+) in the non-clinical sub-sample (HC) and based on the PANSS-P1 item [@Kay1987] (P1<3 for D- and P1>4 for D+) in the clinical sub-sample (SZ).The samples differed in age, with the SZ groups being significantly older than the HC groups, (F(`r aov.age$anova_table$'num Df'[1]`, `r aov.age$anova_table$'den Df'[1]`) = `r aov.age$anova_table$F[1]`, p = `r aov.age$anova_table$'Pr(>F)'[1]`) as shown in Table 1.

#Behavioral Task
Participants performed 160 trials of a probabilistic reversal-learning task with stable and volatile task phases [@Boehme2015] [@Deserno2020]. Participants chose between two card stimuli that were presented in random ordering on the right and left side on the screen. Reward contingencies were anticorrelated (80/20). In the first 55 trials, the more rewarding card remained stable, followed by 55 trials of a volatile task phase, where contingencies changed every 15 to 20 trials. The last block consisted of 35 stable trials with a reversal of the more rewarding card from the first phase. The exact task structure was unique for  every participant.
Behavioral Analysis
Statistical analysis was performed using R software (R Core Team, 2020) and Matlab (). Raw behavioral choice data was analyzed regarding accuracy. To this end, we performed a 2x2x3 mixed factor ANOVA with the between-factors Delusion and Illness and the within-factor phase on the percentage of correct responses, defined as choosing the currently rewarded card.

#Computational Modelling
Model fitting was performed using the HGF toolbox v5.1 as part of the TNU Algorithms for Psychiatry-Advancing Science (TAPAS, http://www.translationalneuromodeling.org/tapas/). Model selection was performed with random-effects Bayesian Model Selection [BMS, spm_BMS in SPM12, ww.fil.ion.ucl.uk/spm] to compute free energy per subject and model, which allowed us to identify the best model according to protected exceedance probability (pxp). Pxp quantifies the likelihood that a model is more frequent than any of the models compared to and is protective against the “null” hypothesis that models are equally frequent within the population (Rigoux, Stephan, Friston, & Daunizeau, 2014). We applied the same model space as (Cole et al., 2020) to individual trial-by-trial choice data of participants. This model space encompasses a i) Rescorla Wagner Learning model, ii) 2-level Hierarchical Gaussian Filter model (HGF2), iii) standard 3-level HGF and iv) 3-level ‘mean-reverting’ HGF with an additional parameter representing constant subject-specific equilibria that volatility estimates remain attracted to. 
The Rescorla-Wagner (Rescorla & Wagner, 1972) model updates Q- action values of a stimulus according to the observed reward prediction error (δ) scaled by a subject-specific learning rate (α).

δQa, t= rt−Qa,tδQa, t= rt−Qa,t

Qa, t+1=Qa,t+ α∗δQa, t

The HGF-models formalize learning as a hierarchical multilevel process in which beliefs about hidden states dynamically evolve in form of Gaussian random walks. The first level (x1) represents whether a respective option was rewarded u(1), which is expressed as a continuous estimate of the reward-probability u(2) on the second level (x2). The third level (x3) describes the volatility of this value u(3). The speed of the evolution of these beliefs (step-size of random walks) is determined by the coupling to the next-highest level (σ1, σ2) and also changes throughout the trial, except on the top (3rd) level representing the constant meta-volatility (ϑ). The second-level belief has a tonic ω and a phasic component κ. The “mean-reverting” model introduced by Cole et al. (2020) introduces another constant component on the third level – the “equilibrium parameter” (m3). Higher values of m3 indicate higher average-levels of the volatility estimates and are in line with hypothesized lower precision of beliefs and higher uncertainty.
Following Bayesian model selection, further analysis was based on the best-fitting model. Priors were set according to the behavioral modelling code in the repository from Cole (2020) to ensure comparability of modelling results. We compared free maximum a posteriori estimates (m3, μ3 k= (0), κ, ω, ϑ, β) across the 4 groups using Type III 2x2 ANOVAs. Finally, we performed a 2x2x3 mixed factor ANOVA, with factors Illness, Delusion and phase (stable, reversal, volatile) on phase-means of trajectories of second-level belief-uncertainty (σ2), third-level volatility (μ3) and precision-weighted PEs of the outcome and volatility (ɛ2, ɛ3).
#fMRI Analysis
Functional Magnetic Resonance Imaging.
fMRI data was acquired on a 3 Tesla Siemens Trio scanner. The protocol included individual field maps acquired before the task to account for individual homogeneity differences of the magnetic field, T1-weighted structural images (TR=1300ms, TE=3.46ms, flip=10°, matrix=240×256, voxel size: 1×1×1mm, slices=170) and gradient echo T2*-weighted echo-planar images with blood oxygenation level dependent contrast (36 slices, covering the whole brain) in oblique orientation at 20° to AC-PC line with 2.5-mm thickness, 3x3mm² in-plane voxel resolution, 0.5- mm gap between slices, TR=2s, TE=22ms and a flip angle α=90°).

#Preprocessing
Neuroimaging data was analyzed with statistic parametric mapping (SPMXXX for preprocessing and SPM12 for first-level and group-level analyses; Wellcome Department of Imaging Neuroscience, London, UK; http://www.fil.ion.ucl.ac.uk/spmhttp://www.fil.ion.ucl.ac.uk/spm/software/spm) in Matlab R2019b (The MathWorks). Images were slice-time and motion-corrected, coregistered, spatially normalized to MNI space using subject-specific T1 scans and smoothed with an isotopic Gaussian kernel of 6mm.

#First-level
On the single-subject-level, we conducted an event-related analysis with the general linear model approach implemented in SPM12. As in previous studies (Iglesias, 2013, Katthagen, 2020) the base regressor spanned from trial onset to end. From the fitted learning trajectories we entered four parametric modulators that changed with onset of the cue, including the outcome prediction (μ1), 1st-level uncertainty (σ1), volatility estimate (μ3) and 2nd level uncertainty (σ2). We entered two modulators that only modelled the feedback phase and encoded the precision weighted prediction errors on the 2nd and 3rd level (ε2, ε3) resulting in six parametric modulators total. Additionally, six movement regressor were included as regressors of no interest as well as their temporal derivates and one regressor to flag scan-to-scan movement >1mm.
Second-level
We used the flexible factorial module of SPM12 to compute 2x2x6x82 ANOVAs, with factors illness, delusion, regressor and subject with age and gender entered as covariates. Linear contrasts were set to examine main and interaction effects of illness and delusion on μ3, ε2 and ε3.

#Results
##Raw Behavioral Results


### Raw Behavioral Results
```{r Descriptives, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE }
# Descriptives of demographics and free model parameters
options(digits = 3)
behav <- group_by(DD_data_fitter, group_DD) %>%
  summarise_at(vars(CorrectPerc, p_stay, Alter), funs(mean, sd))
params <- group_by(DD_data_fitter, cat_group) %>%
   summarise_at(vars(Alter, CorrectPerc,M6_kappa, M6_omega, M6_theta, M6_beta, M6_m3, M6_mu_0), funs(mean, sd))

# make table
t(params) %>%
  kbl(caption = "Summary of demographic and modeling parameters") %>%
  kable_classic(full_width = F, html_font = "Times New Roman")

#Age ANOVA
 aov.age<-DD_data_fitter %>% 
  aov_ez(
    data = .,
    dv = "Alter",
    id = "Code",
    between = c("Illness", "Delusion"))
  
#compare correct responses across groups and phases
mm.corr_perc<- DD_data_long_fitter %>% gather(phase, correct, Correct_pre:Correct_post) %>% 
  aov_ez(
     data = .,
     dv = "correct",
     id = "Code",
     within="phase",
     between = c("fIllness", "fDelusion")  )


# examine pos-hoc tests
correct_phase<- emmeans(mm.corr_perc, c("phase"))
correct_Illness<- emmeans(mm.corr_perc, c("fIllness"))
ref_df <- as.data.frame(summary(correct_phase))
contrast(correct_Illness, method="pairwise") # uses tukey adjustment for multiple testing
contrast(correct_phase, method="pairwise")

#explore correlations between parameters
chart.corrs_new <- chart.Correlation(select(DD_data_fitter,M6_kappa:M6_beta), histogram=TRUE, pch=19)
```

Descriptive and test-statistics of percentage of correct responses and tendency to repeat the choice of the same are shown in Table 1. Across all phases, healthy participants chose more often the correct card compared to patients (t(1, `c_phase@dfargs[["Vdf"]][["Code:phase"]]`, `r mean(DD_data$Alter)` ± `r sd(DD_data$Alter)`y, `r sum(DD_data$Geschlecht == "w")` females) com

```{r one-way_ANOVA, echo=FALSE, warning=FALSE, message= FALSE, error = FALSE}
#2.	One-way ANOVA between CHR and HC on all parameters; Cole: a.	 μ^3 is higher in CHR

# Do the ANOVAs on all model parameters
formulae <- lapply(colnames(DD_data_fitter)[26:32], function(x) as.formula(paste0("dv= '",x,"'")))
ANOVA_one <- lapply(formulae, function(x) afex::aov_ez(data = DD_data_fitter, x, id = "Code", between = c("Illness", "Delusion"))) #covariate = "Alter", factorize = 0))
names(ANOVA_one) <- format(formulae)
F_val <-lapply(names(ANOVA_one), function(x){ANOVA_one[[x]][["anova_table"]][["F"]]})

# -----------------------------following section was to check anova assumptions
# Check ANOVA assumptions; spread of all groups in plot 1 and 3 similar?
# plot linear model from ANOVA
par(mfrow=c(2,2))
plot(ANOVA_one[[1]][["lm"]]) # do that for all ANOVAs an enter index manually (1-7)
# Shapiro to check normality
Alter <-array(DD_data_fitter$Alter)
vars_to_check <- DD_data_fitter[26:32]
apply(vars_to_check, MARGIN = 2,  FUN = function(x) shapiro.test(x))
# Outliers: check number of extreme values (>3SD)
apply(vars_to_check, MARGIN = 2,  FUN = function(x) sum(is_extreme(x)))
# -> yes, many have outliers and omega, theta have extreme values (sigma not analyzed)
#omega - not so bad
o <- print(identify_outliers(as.data.frame(DD_data$M6_omega)))
plot(DD_data$M6_omega)
outliervals = o$`DD_data$M6_omega`[o$is.extreme]
DD_data_omex <- DD_data[DD_data$M6_omega!=outliervals[1] & DD_data$M6_omega!=outliervals[2], ]
#theta
th <- identify_outliers(as.data.frame(DD_data$M6_theta))
plot(DD_data$M6_theta)
outliervals = o$`DD_data$M6_omega`[o$is.extreme]
DD_data_thex <- DD_data[DD_data$M6_theta!=outliervals[1] & DD_data$M6_theta!=outliervals[2], ]
#2. Levenes test for sphericity, omega is not homoscedastic
#apply(vars_to_check, MARGIN = 2,  FUN = function(x) leveneTest(residuals(ANOVA_one$x), center = mean))
# Cooks Distance larger than 4/82?
sum(cooks.distance(ANOVA_one$M6_beta$lm)>.049)#mean(DD_data$M6_beta)*3)
plot(ANOVA_one$M6_beta$lm,4)
#plot residuals
PlotQQ(residuals(ANOVA_one$M6_beta))
#check normality of residuals
#ols_test_normality(residuals(ANOVA_one$M6_beta))

#---------------------------------------- this is all not great, therefore --> robust anova
vars_to_check <- DD_data_fitter[26:32]
Anova_one_robust <-apply(vars_to_check, MARGIN = 2,  FUN = function(x) raov(x~fIllness*fDelusion,data=DD_data_fitter))
F_val <-lapply(names(Anova_one_robust), function(x){Anova_one_robust[[x]][["table"]][1:3,4:5]})
F_val <- data.frame(F_val)
F_val.new = F_val[seq(1, ncol(F_val), 2), ]
t(data.frame(F_val.new)) %>%
  kbl(caption = "Summary of demographic and modeling parameters") %>%
  kable_classic(full_width = F, html_font = "Times New Roman")

summary(oneway.rfit(DD_data_fitter$M6_m3, DD_data_fitter$group_DD))

# Significant Illness effects on beta and m3 --> post-hoc test
m3_posthoc<- emmeans(Anova_one_robust$M6_m3, c("fIllness"))
contrast(m3_posthoc)

beta_posthoc<- emmeans(Anova_one_robust$M6_beta, c("fIllness"))
contrast(beta_posthoc)

theta_posthoc<- emmeans(ANOVA_one$M6_theta, c("Delusion"))
contrast(theta_posthoc)

```

```{r sa2_phase,  echo=FALSE, warning=FALSE,  message= FALSE, error = FALSE}
histogram(DD_data_fitter$sa2)

mm.sa2_phase<- DD_data_long_fitter%>%
  gather(phase, sa2_traj, sa2_pre:sa2_post) %>%
  afex::aov_ez(
     data = .,
     dv = "sa2_traj",
     id = "Code",
     within="phase",
     between = c("Illness", "Delusion"))

shapiro.test(residuals(mm.sa2_phase))

# mm suggests significant phase and Illness effect - post hoc tests:
sa2phase<- emmeans(mm.sa2_phase, c("phase"))
pairs(sa2phase)
sa2Illness<- emmeans(mm.sa2_phase, c("Illness"))
pairs(sa2Illness)
sa2Delusion<- emmeans(mm.sa2_phase, c("Delusion"))
pairs(sa2Delusion)

```

```{r mu3_phase, echo=FALSE, warning=FALSE,  message= FALSE, error = FALSE}
# mixed model Illness x Delusion x Phase for mu3(k)
mm.mu3k_phase<- DD_data_long_fitter%>% gather(phase, mu3_traj, mu3_pre:mu3_post) %>% 
  afex::aov_ez(
     data = .,
     dv = "mu3_traj",
     id = "Code",
     within="phase",
     between = c("Illness", "Delusion")  ) 

# mm suggests illness and phase effects - post-hoc:
mu3kphase<- emmeans(mm.mu3k_phase, c("phase"))
pairs(mu3kphase)
mu3Illness<- emmeans(mm.mu3k_phase, c("Illness"))
pairs(mu3Illness)
```

```{r epsi2_phase, echo=FALSE, warning=FALSE,  message= FALSE, error = FALSE}
mm.epsi2_phase<- DD_data_long_fitter %>%gather(phase, epsi2_traj, epsi2_pre:epsi2_post) %>%
  afex::aov_ez(
     data = .,
     dv = "epsi2_traj",
     id = "Code",
     within="phase",
     between = c("Illness", "Delusion")  )

e2illphase<- emmeans(mm.epsi2_phase, c("phase", "Delusion"))
pairs(e2illphase)
```

```{r epsi3_phase, echo=FALSE, warning=FALSE,  message= FALSE, error = FALSE}
mm.epsi3_phase<- DD_data_long_fitter %>%gather(phase, epsi3_traj, epsi3_pre:epsi3_post) %>%
  afex::aov_ez(
     data = .,
     dv = "epsi3_traj",
     id = "Code",
     within="phase",
     between = c("Illness", "Delusion")  )
```


```{r yeu_phase, echo=FALSE, warning=FALSE,  message= FALSE, error = FALSE}
mm.yeu_phase<- DD_data_long_fitter %>%gather(phase, yeu_traj, yeu_pre:yeu_post) %>%
  afex::aov_ez(
     data = .,
     dv = "yeu_traj",
     id = "Code",
     within="phase",
     between = c("Illness", "Delusion")  )

```